---
title: "R Notebook"
output: html_notebook
author:
- Ghanem Abdelghani
- Ouifak Hafsaa
subtitle: "Social network analysis project"
output: html_notebook
---

Needed packages :

```{r}
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jdk-11.0.7') # for 64-bit version
library(xlsx)
library(igraph)
library(plotly)
library(stringr)
library (haven)
library (ggplot2)
```

Reading processed data from excel file
```{r}
squared_data<-read.csv2(file="..\\data\\squared_numeric.csv",header=TRUE)
squared_data
```



Creating a sub-graph from the first region.

```{r}
sample_size=240
sample_data=squared_data[1:sample_size,1:sample_size]
G=graph_from_adjacency_matrix(as.matrix(sample_data),mode=c("directed"),weighted=TRUE)
```

Elimination of non connected components
```{r}
comp = components(G, mode="strong")
comp$no
comp$csize
BigComp = which.max(comp$csize)
Main_graph = induced_subgraph(G, 
    which(comp$membership == BigComp))
G2=Main_graph
```

Define layout and colors

```{r}
L <- layout.sphere(G)
V(G)$color <- "blue"
V(G)[str_detect(V(G)$name,"_R1")]$color <- "Sectors of region 1"  
V(G)[str_detect(V(G)$name,"_R2")]$color <- "Sectors of region 2" 
V(G)[str_detect(V(G)$name,"_R3")]$color <- "Sectors of region 3" 
V(G)[str_detect(V(G)$name,"_R4")]$color <- "Sectors of region 4"
V(G)[str_detect(V(G)$name,"_R5")]$color <- "Sectors of region 5"
V(G)[str_detect(V(G)$name,"_R6")]$color <- "Sectors of region 6"
V(G)[str_detect(V(G)$name,"_R7")]$color <- "Sectors of region 7"
V(G)[str_detect(V(G)$name,"_R8")]$color <- "Sectors of region 8"
V(G)[str_detect(V(G)$name,"_R9")]$color <- "Sectors of region 9"
V(G)[str_detect(V(G)$name,"_R10")]$color <- "Sectors of region 10"
V(G)[str_detect(V(G)$name,"_R11")]$color <- "Sectors of region 11"
V(G)[str_detect(V(G)$name,"_R12")]$color <- "Sectors of region 12"
G_label<-get.vertex.attribute(G)
G_color<-V(G)$color
```

```{r}
L2 <- layout.sphere(G2)
V(G2)$color <- "blue"
V(G2)[str_detect(V(G2)$name,"_R1")]$color <- "Sectors of region 1"  
V(G2)[str_detect(V(G2)$name,"_R2")]$color <- "Sectors of region 2" 
V(G2)[str_detect(V(G2)$name,"_R3")]$color <- "Sectors of region 3" 
V(G2)[str_detect(V(G2)$name,"_R4")]$color <- "Sectors of region 4"
V(G2)[str_detect(V(G2)$name,"_R5")]$color <- "Sectors of region 5"
V(G2)[str_detect(V(G2)$name,"_R6")]$color <- "Sectors of region 6"
V(G2)[str_detect(V(G2)$name,"_R7")]$color <- "Sectors of region 7"
V(G2)[str_detect(V(G2)$name,"_R8")]$color <- "Sectors of region 8"
V(G2)[str_detect(V(G2)$name,"_R9")]$color <- "Sectors of region 9"
V(G2)[str_detect(V(G2)$name,"_R10")]$color <- "Sectors of region 10"
V(G2)[str_detect(V(G2)$name,"_R11")]$color <- "Sectors of region 11"
V(G2)[str_detect(V(G2)$name,"_R12")]$color <- "Sectors of region 12"
G2_label<-get.vertex.attribute(G2)
G2_color<-V(G2)$color
```


Density of the network

```{r}
    #Density
    graph.density(G,loop=TRUE)
    #Average Path Length
    mean_distance(G)
```

```{r}
#Clustering coefficient
transitivity(G2, type="weighted", weights= E(G2)$weight)
```
Centrality measures :

InStrength:

```{r}
    Instrength=strength(G, vids = V(G), mode = "in",loops = TRUE, weights = E(G)$weight)
    Instrength=as.data.frame(Instrength)
    #G_color=V(G)$color
    #qplot(DegreeDis, data=DegreeDis2, geom="histogram", binwidth=.01)
    fig_instr <- plot_ly(
      x = c(1:length(V(G)$name)),
      y = Instrength$Instrength,
      text=~paste(paste('Sector: ',G_label$name)),
      #name = "InDegree distribution of industries of ",
      #type = "bar",
      color = ~G_color
    )
   
    fig_instr <- fig_instr %>% layout(
         xaxis = list(title = ""),
         yaxis = list(title = ""))
    fig_instr <- fig_instr%>%add_markers(showlegend=F)
    fig_instr
```

OutStrength:

```{r}
    Outstrength=strength(G, vids = V(G), mode = "out",loops = TRUE, weights = E(G)$weight)
    Outstrength=as.data.frame(Outstrength)
    #qplot(DegreeDis, data=DegreeDis2, geom="histogram", binwidth=.01)
    fig_outstr <- plot_ly(
      x = c(1:length(V(G)$name)),
      y = Outstrength$Outstrength,
      text=~paste(paste('Sector: ',G_label$name)),
      #name = "In/Out strength of industries of region 1",
      #type = "bar",
      color = ~G_color
    )
    fig_outstr <- fig_outstr %>% layout(
         xaxis = list(title = ""),
         yaxis = list(title = "strength"))
    fig_outstr <- fig_outstr%>%add_markers(showlegend=F)
    fig_outstr
    fig_str=subplot(fig_instr,fig_outstr)
    fig_str
```

InCloseness:

```{r}
#Closeness Centrality
    #Closeness Centrality
    clos=closeness(G2, mode = "in",weights = E(G2)$weight)
    inclos=as.data.frame(clos)
    fig_inclos <- plot_ly(
      x = c(1:length(V(G2)$name)),
      y = inclos$clos,
      text=~paste(paste('Sector: ',G2_label$name)),
      color = ~G2_color
    )
    fig_inclos <- fig_inclos %>% layout(#title = "InCloseness of industries of Morocco",
         xaxis = list(title = ""),
         yaxis = list(title = "InCloseness"))
    fig_inclos <- fig_inclos%>%add_markers(showlegend=F)

    fig_inclos
```

OutCloseness: 

```{r}
#Closeness Centrality
    #Closeness Centrality
    clos=closeness(G2, mode = "out",weights = E(G2)$weight)
    outclos=as.data.frame(clos)
    fig_outclos <- plot_ly(
      x = c(1:length(V(G2)$name)),
      y = outclos$clos,
      text=~paste(paste('Sector: ',G2_label$name)),
      color = ~G2_color
    )
    fig_outclos <- fig_outclos %>% layout(#title = "OutCloseness of industries of Morocco",
         xaxis = list(title = ""),
         yaxis = list(title = "OutCloseness"))
    fig_outclos <- fig_outclos%>%add_markers(showlegend=F)

    fig_outclos
```

Subplot of the two figures of closeness:

```{r}
fig_clos <- subplot(fig_inclos, fig_outclos) 
fig_clos <- fig_clos %>% layout(title = "In/out Closeness of industries of Morocco",
         xaxis = list(title = ""),
         yaxis = list(title = "Closeness"))
fig_clos
```

AllCloseness:


```{r}
    clos=closeness(G2, mode = "all")
    allclos= as.data.frame(clos)
    fig22 <- plot_ly(
      x = c(1:length(V(G2)$name)),
      y =allclos$clos,
      text=~paste(paste('Sector: ',G2_label$name)),
      color = ~G2_color
    )
    fig22 <- fig22 %>% layout(title = "AllCloseness of industries of Morocco",
         xaxis = list(title = ""),
         yaxis = list(title = "MAD"))
    
    fig22
```

Betweeness:


```{r}
#G_label
#Betweeness Centrality
    Betweeness <- betweenness(G,weights=E(G)$weight, directed=TRUE,nobigint=TRUE)
    Betweeness <- as.data.frame(Betweeness)
    fig_bet=plot_ly( x = c(1:length(V(G)$name)), y = Betweeness$Betweeness,text=~paste(paste('Sector: ',G_label$name)),color = ~G_color)
    fig_bet <- fig_bet %>% layout(#title = "Betweeness of industries of Morocco",
         xaxis = list(title = ""),
         yaxis = list(title = "Betweeness"),
         width = 800, 
          height = 500)
        fig_bet <- fig_bet%>%add_markers(showlegend=F)
    fig_bet
```

EigenCentrality:                                                                 
                                                                 
```{r}
EigenCentrality <- eigen_centrality(G, directed = TRUE, weights = E(G)$weight)
   EigenCentrality <- as.data.frame(EigenCentrality)
   fig_eig <- plot_ly(
        x = c(1:length(V(G)$name)),
        y = EigenCentrality$vector,
        text=~paste(paste('Sector: ',G_label$name)),
        #type = "bar",
        color = ~G_color
       )
      fig_eig <- fig_eig %>% layout(title = "EigenCentrality of industries of Morocco",
           xaxis = list(title = ""),
           yaxis = list(title = "Eigenvector Centrality"),
           width = 800, 
           height = 500)
      #fig_eig <- fig_eig%>%add_markers(showlegend=F)
      fig_eig
```

PageRank:   
                                                                    
```{r}
PageRank <- page_rank(G, directed = TRUE, weights = E(G)$weight)
#PageRank = as.data.frame(PageRank)
   fig_pr <- plot_ly(
        x = c(1:length(V(G)$name)),
        y = PageRank$vector,
        text=~paste(paste('Sector: ',G_label$name)),
        #type = "bar",
        color = ~G_color
       )
      fig_pr <- fig_pr %>% layout(title = "Page rank measure of industries of Morocco",
           xaxis = list(title = ""),
           yaxis = list(title = "Page Rank Centrality"),
           width = 800, 
           height = 500)
      fig_pr
```
 
 Subplot of all figures:  
   
```{r}
# subplot
fig <- subplot(fig_instr,fig_outstr,fig_inclos,fig_outclos, fig_bet, fig_eig,fig_pr,nrows = 4) 
fig  %>% layout(title="Centrality measures of Moroccan sectors network",
  width = 760, 
  height = 650,
  annotations = list(
    list(x = 0.2 , y = 0.95, text = "In Strength", showarrow = F, xref='paper', yref='paper'),
    list(x = 0.8 , y = 0.95, text = "Out Strength", showarrow = F, xref='paper', yref='paper'),
    list(x = 0.2 , y = 0.7, text = "In closeness", showarrow = F, xref='paper', yref='paper'),
 list(x = 0.8 , y = 0.7, text = "Out closeness", showarrow = F, xref='paper', yref='paper'),
  list(x = 0.2 , y = 0.45, text = "Betweeness", showarrow = F, xref='paper', yref='paper'),
 list(x = 0.8 , y = 0.45, text = "EigenVector", showarrow = F, xref='paper', yref='paper'),
 list(x = 0.2 , y = 0.16, text = "Page Rank", showarrow = F, xref='paper', yref='paper'))
)
```
 
 Reading the output multiplier data and convert it to numeric
 
```{r}
data2<-read.xlsx2("..\\data\\output_mul.xlsx",header=FALSE,sheetName = "B_2")
data2=as.data.frame(data2)
data2[] <- lapply(data2, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(data2, class)
data2[is.na(data2)]=0
data2=as.vector(t(data2))
```
 
 
 Testing the correlations between eigen vector centrality and output multiplier                                                                                            
```{r}
vec1=EigenCentrality$vector
vec2=PageRank$vector
vec3=Betweeness$Betweeness
vec4=Instrength$Instrength
cor.test(data2,vec1 ,method="spearman")
cor(data2,vec4)
cor(data2,vec4)
cor(data2,vec4)
cor(data2,vec4)
```

Calculating closeness for even non connected components
```{r}
clos=closeness(G, mode = "all",weights = E(G)$weight)
allclos= as.data.frame(clos)
clos=closeness(G, mode = "in",weights = E(G)$weight)
inclos= as.data.frame(clos)
clos=closeness(G, mode = "out",weights = E(G)$weight)
outclos= as.data.frame(clos)
```

Calculating the correlations betwen the different centrality measures and the output multiplier :

```{r}
EigenVector=EigenCentrality$vector
Betweeness=Betweeness$Betweeness
In_Closeness=inclos$clos
Out_Closeness=outclos$clos
All_Closeness=allclos$clos
In_Strength=Instrength$Instrength
Out_Strength=Outstrength$Outstrength
PageRank=PageRank$vector
output=data2
dataf=data.frame(In_Strength,Out_Strength,In_Closeness,Out_Closeness,Betweeness,EigenVector,PageRank,output)
corm=cor(dataf)
library(corrplot)
corrplot(corm, method="color")
```
                                                                                             