---
title: "Morocco's MRIO network 2013"
subtitle: "Social network analysis project"
author:
- Ghanem Abdelghani
- Ouifak Hafsaa
output: html_notebook
---





Including the necessary packages
```{r}
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jdk-11.0.7') # for 64-bit version
library(xlsx)
library(igraph)
library(plotly)
library(stringr)
```

Reading processed data from excel file
```{r}
squared_data<-read.xlsx2("..\\data\\processed_morocco_MRIO.xlsx",header=TRUE,sheetName = "IIOS")

squared_data
```

Converting the data to numeric and save it in an external file
```{r}
squared_data[] <- lapply(squared_data, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(squared_data, class)

squared_data[is.na(squared_data)]=0
  write.table(squared_data, file = "..\\data\\squared_numeric.csv", sep = ";",row.names = FALSE)


```

Creating a sub-graph from the first 6 regions, as it's too large to plot the entire network.
Also we should note that the visualization of the nework is not going to be that much of a help. since the network is dense ( almost complete ) which makes it nearly impossible to derive any insights from it. 
```{r}
sample_size=120
sample_data=squared_data[1:sample_size,1:sample_size]
G3=graph_from_adjacency_matrix(as.matrix(sample_data),mode=c("directed"),weighted=TRUE)
```

Defining the layout and the colors of nodes
```{r}
L <- layout.sphere(G3)


V(G3)$color <- "blue"
V(G3)[str_detect(V(G3)$name,"_R1")]$color <- "Sectors of region 1"  
V(G3)[str_detect(V(G3)$name,"_R2")]$color <- "Sectors of region 2" 
V(G3)[str_detect(V(G3)$name,"_R3")]$color <- "Sectors of region 3" 
V(G3)[str_detect(V(G3)$name,"_R4")]$color <- "Sectors of region 4"
V(G3)[str_detect(V(G3)$name,"_R5")]$color <- "Sectors of region 5"
V(G3)[str_detect(V(G3)$name,"_R6")]$color <- "Sectors of region 6"
```

Elemenating multiple links between two nodes for visualization purposes
```{r}
G<-simplify(G3, remove.multiple = TRUE, remove.loops = TRUE,edge.attr.comb = igraph_opt("sum"))
```

Setting up coordinates,atributes, colors and sizes of the nodes
```{r}
vs <- V(G)
es <- as.data.frame(get.edgelist(G))

Nv <- length(vs)
Ne <- length(es[1]$V1)

Xn <- L[,1]
Yn <- L[,2]
G_label<-get.vertex.attribute(G)
G_color<-V(G)$color
total<-c()
node_size<-c()
 for (i in 1:ncol(squared_data[,1:sample_size])){
   total[i]<-sum(squared_data[1:nrow(squared_data),i])
   node_size[i]<-sum(squared_data[1:nrow(squared_data),i])
 }

node_size <-100*(node_size-min(node_size))/(max(node_size)-min(node_size))

network <- plot_ly(x = ~Xn, y = ~Yn, mode = "markers", text=~paste(paste('Sector: ',G_label$name),'<br />',sep='',paste('Total: ',total,'Million MAD')) ,color = ~G_color,marker = list(size = ~node_size, opacity = 0.9) , hoverinfo = "text")
```

Setting up the edges
```{r}
edge_shapes <- list()
for(i in 1:Ne) {
  v0 <- es[i,]$V1
  v1 <- es[i,]$V2
  
  edge_shape = list(
    type = "line",
    line = list(color = "#030303", width = 0.02),
    x0 = Xn[v0],
    y0 = Yn[v0],
    x1 = Xn[v1],
    y1 = Yn[v1]
  )

  edge_shapes[[i]] <- edge_shape
}
```

Setting up the figure's parametrs
```{r}
axis <- list(title = "", showgrid = FALSE, showticklabels = FALSE, zeroline = FALSE)

fig <- layout(
  network,
  title = 'Sample of the Interregional input-output network of morocco, 2013',
  shapes = edge_shapes,
  autosize = F,
  width = 1000, 
  height = 1000, 
  xaxis = axis,
  yaxis = axis
)
```


List of Regions
---
Symbol  Region
------  -------
R1	    Tanger-Tetouan-Al Hoceima
R2	    Oriental
R3	    Fès-Meknès
R4	    Rabat-Salé-Kénitra
R5	    Béni Mellal-Khénifra
R6	    Grand Casablanca-Settat
R7	    Marrakech-Safi
R8	    Drâa-Tafilalet
R9	    Souss-Massa
R10	    Guelmim-Oued Noun
R11	    Laayoune-Sakia El Hamra
R12	    Dakhla-Oued Eddahab
---


List of sectors
---
Symbol  French Name                                         English Name
------  -----------                                         -------------
A00	    Agriculture, forêt et services annexes	            Agriculture, forestry, hunting, related services
B05	    Pêche, aquaculture	                                Fishing, aquaculture
C00	    Industrie d’extraction	                            Mining industry
D01	    Industries alimentaires et tabac	                  Food industry and tobacco
D02	    Industries du textile et du cuir	                  Textile and leather industry
D03	    Industrie chimique et parachimique	                Chemical and para-chemical industry
D04	    Industrie mécanique, métallurgique et électrique	  Mechanical, metallurgical and electrical industry
D05	    Autres industries manufac. hors raffinage pétrole	  Other manufacturing, excluding petroleum refining
D06	    Raffinage de pétrole et autres produits d’énergie	  Oil refining and other energy products
E00	    Electricité et eau	                                Electricity and water
F45	    Bâtiment et travaux                                 publics	Construction
G00	    Commerce	                                          Trade
H55	    Hôtels et restaurants	                              Hotels and restaurants
I01	    Transports	                                        Transport
I02	    Postes et télécommunications	                      Post and telecommunications
J00	    Activités financières et assurances	                Financial activities and insurance
K00	    Immobilier, location et serv. rendus entreprises	  Real estate, renting and services to enterprises
L75	    Administration publique et sécurité sociale	        General public administration and social security
MNO	    Education, santé et action sociale	                Education, health and     social                                              action
OP0	    Autres services non financiers	                    Other non-financial services
---


```{r}
fig
    
```


