---
title: "Data preprocessing"
author:
- Ghanem Abdelghani
- Ouifak Hafsaa
subtitle: "Social network analysis project"
output: html_notebook
runtime: shiny
---



Reading data excel file

```{r}
library(xlsx)
library(stringr)
data = read.xlsx2("..\\data\\morocco_MRIO.xlsx",header=TRUE,sheetName = "IIOS")
original_data<-data
```

Here's the initial data
```{r}
original_data
```


First of all, you will notice that we cannot create the graph that we need from this data. In fact we need to create a squared matrix with the node the sectors of each region as nodes plus other nodes (See the article for more details).

Here is what these regions and sectors refere to:

List of Regions
---
Symbol  Region
------  -------
R1	    Tanger-Tetouan-Al Hoceima
R2	    Oriental
R3	    Fès-Meknès
R4	    Rabat-Salé-Kénitra
R5	    Béni Mellal-Khénifra
R6	    Grand Casablanca-Settat
R7	    Marrakech-Safi
R8	    Drâa-Tafilalet
R9	    Souss-Massa
R10	    Guelmim-Oued Noun
R11	    Laayoune-Sakia El Hamra
R12	    Dakhla-Oued Eddahab
---


List of sectors
---
Symbol  French Name                                         English Name
------  -----------                                         -------------
A00	    Agriculture, forêt et services annexes	            Agriculture, forestry, hunting, related services
B05	    Pêche, aquaculture	                                Fishing, aquaculture
C00	    Industrie d’extraction	                            Mining industry
D01	    Industries alimentaires et tabac	                  Food industry and tobacco
D02	    Industries du textile et du cuir	                  Textile and leather industry
D03	    Industrie chimique et parachimique	                Chemical and para-chemical industry
D04	    Industrie mécanique, métallurgique et électrique	  Mechanical, metallurgical and electrical industry
D05	    Autres industries manufac. hors raffinage pétrole	  Other manufacturing, excluding petroleum refining
D06	    Raffinage de pétrole et autres produits d’énergie	  Oil refining and other energy products
E00	    Electricité et eau	                                Electricity and water
F45	    Bâtiment et travaux                                 publics	Construction
G00	    Commerce	                                          Trade
H55	    Hôtels et restaurants	                              Hotels and restaurants
I01	    Transports	                                        Transport
I02	    Postes et télécommunications	                      Post and telecommunications
J00	    Activités financières et assurances	                Financial activities and insurance
K00	    Immobilier, location et serv. rendus entreprises	  Real estate, renting and services to enterprises
L75	    Administration publique et sécurité sociale	        General public administration and social security
MNO	    Education, santé et action sociale	                Education, health and     social                                              action
OP0	    Autres services non financiers	                    Other non-financial services
---


we will start by adding the symbols of the regions to the nodes
```{r}

data<-original_data
for (i in 1:12){
  start=(i-1)*20+1
  end=i*20
  for(j in start:end){
    names(data)[j+1] <- paste(str_remove(names(data)[j+1],"\\.[0-9]*$"),paste("_R",i,sep=""),sep="") 
  }
}
data

```

Naming investement demand as INV_D

```{r}

  start_inv_d=12*20+2
  end_inv_d=start_inv_d+11
  i=1
  for(j in start_inv_d:end_inv_d){
    names(data)[j] <- paste("INV_D_R",i,sep="")
    i=i+1
  }
```

Naming Household demand as HO_D

```{r}

  start_ho_d=end_inv_d+1
  end_ho_d=start_ho_d+11
  i=1
  for(j in start_ho_d:end_ho_d){
    names(data)[j] <- paste("HO_D_R",i,sep="")
    i=i+1
  }
```


Naming Governoment demand as GOV_D
```{r}

  start_gov_d=end_ho_d+1
  end_gov_d=start_gov_d+11
  i=1
  for(j in start_gov_d:end_gov_d){
    names(data)[j] <- paste("GOV_D_R",i,sep="")
    i=i+1
  }
  
```


Naming Foreign exports as FO_EXP
```{r}
end_for_exp<-ncol(data)
names(data)[end_for_exp] <- "FO_EXP"

```


```{r}
data
```



The purpose of this section is to make the data frame squared for graph construction <br/>

Adding the values added

```{r}
end=12*20+1

squared_data=data.frame(data[,2:end])
squared_data
```

We will create theses duplicates so that if we want to re-run the cell code we'll not have to run the entire code from the bigenning (note that the data changes)
```{r}
squared_data_dup=squared_data

```

Here we want to make the matrix squared by adding zeros

Adding foreign investement 
```{r}
squared_data=squared_data_dup
start=ncol(squared_data)+1
end=start+19
for (i in start:end){
  squared_data=cbind(squared_data,colname=numeric(nrow(squared_data)))
  names(squared_data)[ncol(squared_data)] <- paste(data[i,1],"_FOR",sep="")
  
}

squared_data_dup2=squared_data

```

Naming and adding taxes and subsidies, labor payments, capital payments, other costs, value added (resp.) as TAX_SUB, LA_PAY, CA_PAY, OTH_COT, VA_ADD

```{r}
squared_data=squared_data_dup2
for (i in 1:5){
  squared_data=cbind(squared_data,colname=numeric(nrow(squared_data)))
  switch(i,
         names(squared_data)[ncol(squared_data)] <- "TAX_SUB",
         names(squared_data)[ncol(squared_data)] <- "LA_PAY",
         names(squared_data)[ncol(squared_data)] <- "CA_PAY",
         names(squared_data)[ncol(squared_data)] <- "OTH_COT",
         names(squared_data)[ncol(squared_data)] <- "VA_ADD",
         )
}

```

```{r}
squared_data
squared_data_dup3=squared_data
```

```{r}
squared_data=squared_data_dup3
squared_data=cbind(squared_data,data[,start_inv_d:end_for_exp])
```


```{r}
squared_data
```

```{r}
n=302-265
m_zeros=matrix(data=0,nrow=n,ncol = ncol(squared_data))
squared_data=data.frame(rbind(as.matrix(squared_data), as.matrix(m_zeros)))
squared_data
```

Now we will fill the matrix with the actual values of the added columns.

First we will need to convert the data to numeric in order to avoid any casting/levels problems.

Converting data to numeric
```{r}
squared_data[] <- lapply(squared_data, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
sapply(squared_data, class)

squared_data[is.na(squared_data)]=0

squared_data_dup4<-squared_data
```

Filling the matrix with actual values of the added columns
```{r}
squared_data<-squared_data_dup4
start=12*20+1
end=start+24

squared_data[1:ncol(data)-1,start:end]<-as.numeric(t(data[start:end,2:ncol(data)]))

```

```{r}
write.xlsx2(squared_data,"..\\data\\processed_morocco_MRIO.xlsx",sheetName = "IIOS",row.names = FALSE)
```








